---
title: "Data Analysis Final"
author: "Matthew Jusino"
date: "2022-12-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
require(here)
del = read.csv(here("data", "delomys.csv"))
```

```{r}
summary(del$body_mass)
summary(del$body_length)
shapiro.test(del$body_mass)
shapiro.test(del$body_length)
```

```{r}
plot(x = del$body_length, y = del$body_mass)
hist(del$body_length)
hist(del$body_mass)
boxplot(body_mass ~ binomial, data = del)
boxplot(body_mass ~ sex, data = del)
boxplot(body_mass ~ sex:binomial, data = del, las = 2, cex.axis = 0.3, xlab = '')
```

### Q1 (2 pts.): Qualitatively describe the relationship between body mass and length.
### Does the relationship seem linear, curved, nonexistent?

There seems to be a linear relationship between body mass and length.

### Q2 (2 pts.): Qualitatively describe the shapes of the histograms.
### Do the data appear normally-distributed? Explain why or why not.
### Explain why we care (or not) whether the data are normally distributed.

Neither of the histograms appear to be normally distributed. They visibly lack the distinct bell shape of a normal distribution, though body mass is closer. We don't care because the normality assumption for ANOVA applies only to the residuals of the data.

### Q3 (2 pts.): Using both the histograms and normality tests, do you think the (unconditioned) body masses and body length are normally-distributed?

The Shapiro Tests confirm that neither are normally distributed.

### Q4 (2 pts.): Examine the three conditional boxplots.
### Describe any graphical evidence you see for body mass differences based on species and/or sex.

The sublineatus species tend to be slightly lighter, based on the conditional boxplots. There does not appear to be significant difference between sexes.

```{r}
fit1 = lm(body_length ~ body_mass, data = del)
summary(fit1)
fit2 = aov(body_mass ~ sex, data = del)
summary(fit2)
fit3 = aov(body_mass ~ binomial, data = del)
summary(fit3)
fit4 = aov(body_mass ~ sex + binomial, data = del)
summary(fit4)
fit5 = aov(body_mass ~ sex * binomial, data = del)
summary(fit5)
```