---
title: "Data Analysis Final"
author: "Matthew Jusino"
date: "2022-12-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
require(here)
del = read.csv(here("data", "delomys.csv"))
```

```{r}
summary(del$body_mass)
summary(del$body_length)
shapiro.test(del$body_mass)
shapiro.test(del$body_length)
```

```{r}
plot(x = del$body_length, y = del$body_mass)
hist(del$body_length)
hist(del$body_mass)
boxplot(body_mass ~ binomial, data = del)
boxplot(body_mass ~ sex, data = del)
boxplot(body_mass ~ sex:binomial, data = del, las = 2, cex.axis = 0.3, xlab = '')
```

### Q1 (2 pts.): Qualitatively describe the relationship between body mass and length.
### Does the relationship seem linear, curved, nonexistent?

There seems to be a linear relationship between body mass and length.

### Q2 (2 pts.): Qualitatively describe the shapes of the histograms.
### Do the data appear normally-distributed? Explain why or why not.
### Explain why we care (or not) whether the data are normally distributed.

Neither of the histograms appear to be normally distributed. They visibly lack the distinct bell shape of a normal distribution, though body mass is closer. We don't care because the normality assumption for ANOVA applies only to the residuals of the data.

### Q3 (2 pts.): Using both the histograms and normality tests, do you think the (unconditioned) body masses and body length are normally-distributed?

The Shapiro Tests confirm that neither are normally distributed.

### Q4 (2 pts.): Examine the three conditional boxplots.
### Describe any graphical evidence you see for body mass differences based on species and/or sex.

The sublineatus species tend to be slightly lighter, based on the conditional boxplots. There does not appear to be significant difference between sexes.

```{r}
fit1 = lm(body_length ~ body_mass, data = del)
summary(fit1)
shapiro.test(residuals(fit1))
fit2 = aov(body_mass ~ sex, data = del)
summary(fit2)
shapiro.test(residuals(fit2))
fit3 = aov(body_mass ~ binomial, data = del)
summary(fit3)
shapiro.test(residuals(fit3))
fit4 = aov(body_mass ~ sex + binomial, data = del)
summary(fit4)
shapiro.test(residuals(fit4))
fit5 = aov(body_mass ~ sex * binomial, data = del)
summary(fit5)
shapiro.test(residuals(fit5))
```
### Q5 (2 pts.): What do you conclude about residual normality based on the numerical and graphical diagnostics?
Based on the numerical and graphical diagnostics, the residuals do not appear to meet the assumption of normality for any of the models

### Q6 (1 pt.): Are violations of the normality assumption equally severe for all the models?
The violations are severe for all of the models. The closest to the threshold p-value of 0.05 is fit2 with a p-value of 0.0001541.

```{r}
knitr::kable(coef(summary(fit1)))
```

### Q7 (2 pts.): What is the magnitude of the mass/length relationship?

The magnitude of the mass/length relationship is 0.875

### Q8 (2 pts.): What is the expected body length of an animal that weighs 100g?
The expected body length for an animal weighing 100g is 114mm.

### Q9 (2 pts.): What is the expected body length of an animal that weighs 0g?
The expected body length for an animal weighing 0g is 0mm

```{r}
knitr::kable(coef(fit2))
knitr::kable(coef(fit3))
knitr::kable(coef(fit4))
knitr::kable(coef(fit5))
```

### Q10 (1 pt.): What is the base level for sex?
The base level for sex is female

### Q11 (1 pt.): What is the base level for binomial?
The base level for binomial is D.dorsalis

### Q12 (1 pt.): Which sex is heavier? How do you know?
Males are heavier, as their coefficient is positive, meaning their mass value is greater than the base case (female).

### Q13 (1 pt.): Which species is heavier? How do you know?
D.dorsalis is heavier, as the D.sublineatus coefficient is negative, meaning their mass value is less than the base case (D.Dorsalis).

```{r}
knitr::kable(anova(fit2))
knitr::kable(anova(fit3))
knitr::kable(anova(fit4))
knitr::kable(anova(fit5))
```

### Q14 (1 pt.): Are sex and species significant predictors for body mass?
Both are significant predictors for body mass

### Q15 (1 pt.): Is there a significant interaction?
There is not a significant interaction between the two, as the F-value is very small.

### Q16 (2 pts.): Examine the p-values for the main effects (sex and species) in all four of the ANOVA tables. Does the significance level of either main effect change very much among the different models?
The significance level of either main effect does not change significantly among the different models. Only minor changes on the order of 0.00008

```{r}
AIC(fit1)
AIC(fit2)
AIC(fit3)
AIC(fit4)
AIC(fit5)
```
### Q17 (1 pt.): Which two models have the lowest AIC?


### Q18 (4 pts.): Which of the two models with lowest AIC scores would you select?
### Explain your decision based on model fit and the complexity/understanding tradeoff.
